Data Exploration: IUCN and AquaMaps

### 2016-08-23 clip_depth/coral_maps.Rmd
messing around with the AquaMaps FAO clipping issue.  Plots and maps for quadrant 3 and 4 species...

if(!file.exists(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))) {

  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'coralmaps/ne_10m_bathymetry_K_200'), 
                       layer = 'ne_10m_bathymetry_K_200', 
                       stringsAsFactors = TRUE)

@@

  writeRaster(bath_raster, file.path(dir_git, 'coralmaps', 'bathy_cells.tif'))

  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
  # bath_df <- read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd')
  ### probably want to filter to just 100% 200m+
}

@@

if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))) {

@@

  write_csv(spp_narr, file.path(dir_git, 'coralmaps', 'coral_spp_narratives.csv'))

@@

  write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))

} else 
  spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))

@@

bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
  mutate(deep = 1) %>%
  select(-weight)

spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))

@@

write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))
length(unique(coral_cells_iucn$iucn_sid))
### NOTE: only 473 coral species with IUCN maps
```

``` {r calc_areas_for_iucn_corals}
### calc areas
### read hcaf to get areas; left join to coral_cells_iucn
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
  mutate(deep = ifelse(deep == 1, 'deep', 'shallow'))

@@

write_csv(iucn_coral_areas, file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv'))

@@

``` {r calc_overlaps}
if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))) {

  ### get species list
  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0)

@@

  ### copy aquamaps cells to local file, keeping only corals in the paired maps
  if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))) {
    coral_cells_am <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
      rename(am_sid = speciesid, prob = probability) %>%
      filter(am_sid %in% spp_coralmaps$am_sid) 
    write_csv(coral_cells_am, file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))
  }

  coral_cells_am <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))

@@

  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
}

@@

spp_coral_clip <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv')) %>%
  select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
  mutate(shallow_pct = 100 * shallow_pct,
         deep_pct    = 100 * deep_pct)

@@

spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))

@@

write_csv(iucn_coral_fixed, file.path(dir_git, 'coralmaps', 'iucn_coral_fixed.csv'))

@@

iucn_coral_fixed <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_fixed.csv'))

@@

write_csv(iucn_coral_quads, file.path(dir_git, 'coralmaps', 'iucn_coral_quads.csv'))

@@

iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_quads.csv')) %>%
  select(method, quad, quad_name, pct_quad) %>%
  distinct()
### Plot the bar chart
barchart_coral_quads <- ggplot(iucn_coral_quad_sum, 
                                aes(x = method, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = 1) +
  scale_fill_manual(values = c('q4' = '#d01c8b', 
                               'q3' = '#f1b6da', 
                               'q2' = '#b8e186',
                               'q1' = '#4dac26'),
                    labels = quad_names$quad_name) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = break_nums/100,
                     labels = sprintf('%s%%', break_nums)) + 
  ### add grid lines; horizontal but then get flipped
  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Percent of corals by quadrant', 
       fill = 'Alignment')

@@

write_csv(iucn_coral_quad_sum, file.path(dir_git, 'coralmaps', 'iucn_coral_quad_sum.csv'))

>>>>

if(!file.exists(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'))) {

  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'clip_depth/ne_10m_bathymetry_K_200'), 
                       layer = 'ne_10m_bathymetry_K_200', 
                       stringsAsFactors = TRUE)

@@

  writeRaster(bath_raster, file.path(dir_git, 'clip_depth', 'bathy_cells.tif'))

  write_csv(bath_df, file.path(dir_git, 'clip_depth', 'bathy_cells.csv'))
  # bath_df <- read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd')
  ### probably want to filter to just 100% 200m+
}

@@

if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))) {

@@

  write_csv(spp_narr, file.path(dir_git, 'clip_depth', 'coral_spp_narratives.csv'))

@@

  write_csv(spp_depth, file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))

} else 
  spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))

@@

bath_df <-  read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
  mutate(deep = 1) %>%
  select(-weight)

spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))

@@

write_csv(coral_cells_iucn, file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))
length(unique(coral_cells_iucn$iucn_sid))
### NOTE: only 473 coral species with IUCN maps
```

``` {r calc_areas_for_iucn_corals}
### calc areas
### read hcaf to get areas; left join to coral_cells_iucn
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
coral_cells_iucn <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
  mutate(deep = ifelse(deep == 1, 'deep', 'shallow'))

@@

write_csv(iucn_coral_areas, file.path(dir_git, 'clip_depth', 'iucn_coral_areas.csv'))

@@

``` {r calc_overlaps}
if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))) {

  ### get species list
  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0)

@@

  ### copy aquamaps cells to local file, keeping only corals in the paired maps
  if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))) {
    coral_cells_am <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
      rename(am_sid = speciesid, prob = probability) %>%
      filter(am_sid %in% spp_coralmaps$am_sid) 
    write_csv(coral_cells_am, file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))
  }

  coral_cells_am <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))

@@

  write_csv(spp_coralmaps, file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))
}

@@

spp_coral_clip <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_areas.csv')) %>%
  select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
  mutate(shallow_pct = 100 * shallow_pct,
         deep_pct    = 100 * deep_pct)

@@

spp_coralmaps <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))

@@

write_csv(iucn_coral_fixed, file.path(dir_git, 'clip_depth', 'iucn_coral_fixed.csv'))

@@

iucn_coral_fixed <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_fixed.csv'))

@@

write_csv(iucn_coral_quads, file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv'))

@@

iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv')) %>%
  select(method, quad, quad_name, pct_quad) %>%
  distinct()
### Plot the bar chart
barchart_coral_quads <- ggplot(iucn_coral_quad_sum, 
                                aes(x = method, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = 1) +
  scale_fill_manual(values = c('q4' = '#d01c8b', 
                               'q3' = '#f1b6da', 
                               'q2' = '#b8e186',
                               'q1' = '#4dac26'),
                    labels = quad_names$quad_name) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = break_nums/100,
                     labels = sprintf('%s%%', break_nums)) + 
  ### add grid lines; horizontal but then get flipped
  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Percent of corals by quadrant', 
       fill = 'Alignment')

@@

write_csv(iucn_coral_quad_sum, file.path(dir_git, 'clip_depth', 'iucn_coral_quad_sum.csv'))
